Chemically Self-Consistant Modeling of the Globular Cluster NGC 2808 and its Effects on the Inferred Helium abundance of Multiple Stellar Populations.

### 2023-06-30 src/ms.tex
Finished first pass introduction, split off some stuff into a modeling section
Signifigant progress, wrote most of fidanka section
Transed-Name
Paper updates

\documentclass[twocolumn,linenumbers]{src/aastex631}
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

@@

\correspondingauthor{Thomas M. Boudreaux}
\email{thomas.m.boudreaux.gr@dartmouth.edu, thomas@boudreauxmail.com}

\author[0000-0002-2600-7513]{Thomas M. Boudreaux}
\affiliation{Department of Physics and Astronomy, Dartmouth College, Hanover, NH 03755, USA}

@@

@@

\begin{abstract}
	Over its approximately 100 year history stellar modeling has become an
	essential tool for understanding certain astrophysical phenomena which are not
	directly observable. Modeling allows for empirical constraints --- such as
	elemental abundances, luminosities, and effective temperatures --- to strongly
	inform non-observables such as a star's age, mass, and radius. Here we propose a
	thesis in five parts, related through their use of both modeling and the
	Dartmouth Stellar Evolution Program (DSEP) to conduct this modeling. In two of
	the parts of this thesis we will use DSEP, in conjunction with atmospheric
	boundary conditions generated by collaborators, to build chemically
	self-consistent models of multiple populations (MPs) in the globular clusters
	NGC 2808, 47 Tuc, and NGC 6752. We will infer helium abundances across MPs and
	compare these inferred abundances to those from models which do not consider as
	careful a handling of a star's chemistry. PLACEHOLDER
\end{abstract}

\keywords{}

@@

\begin{acknowledgments}
	This work has made use of the NASA astrophysical data system (ADS). We
	would like to thank Elisabeth Newton and Aaron Dotter for their support and
	for useful disscusion related to the topic of this paper. We acknowledge
	the support of a NASA grant (No. 80NSSC18K0634). 
\end{acknowledgments}

>>>>

\documentclass[twocolumn,linenumbers]{src/aastex631}
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\newcommand\fidanka{\texttt{Fidanka} }

@@

\correspondingauthor{Emily M. Boudreaux}
\email{emily.m.boudreaux.gr@dartmouth.edu, emily@boudreauxmail.com}

\author[0000-0002-2600-7513]{Emily M. Boudreaux}
\affiliation{Department of Physics and Astronomy, Dartmouth College, Hanover, NH 03755, USA}

@@

\author{Rehnata Hoh}
\affiliation{Department of Physics and Astronomy, Dartmouth College, Hanover, NH 03755, USA}

@@

\begin{abstract}
	Globular Clusters (GCs) provide a unique astrophysical laboratory for
	studying the formation and evolution of stars. GCs are old, dense, and it
	has historically been believed that they have a single stellar population
	(SSP, in some publications also simple stellar population). However, in the
	last two decades, it has been definitively shown that most if not all Milky
	Way GCs have multiple stellar populations (MPs). These MPs are chemically
	distinct from one another, primarily separated by light element abundance
	variations without the standard accompanying heavy element abundance
	variations. In addition to the light element abundance variations, MPs also
	have different inferred helium abundances. The origin of these MPs remains
	an open question, and one which is sensitive to the population - population
	compositional differences; consequently, the extent of the composition
	variations is a key parameter to constrain. Metal abundances may be
	directly measured spectroscopically; however, helium abundances are not
	directly observable in GCs. Instead, helium abundances are inferred from
	stellar models. It is therefore important to understand build stellar
	models that are self-consistent in the compositions of the structure,
	atmosphere, and opacity. In this work we present the first chemically
	self-consistent stellar models of the Milky Way Globular Cluster NGC 2808.
	We find that the helium abundance of the second generation of stars is
	higher than the first generation by {\color{red} SOME AMOUNT}.
	{\color{blue} This is in agreement with previous studies of NGC 2808.}
\end{abstract}

\keywords{Globular Clusters (656), Stellar evolutionary models (2046)}

@@

\input{src/sections/AtmopshericModels.tex}

\input{src/sections/modeling.tex}

\input{src/sections/fidanka.tex}

\input{src/sections/fitting.tex}

\input{src/sections/results.tex}

% \input{src/sections/discussion.tex}

\input{src/sections/conclusion.tex}

\begin{acknowledgments}
	This work has made use of the NASA astrophysical data system (ADS). We
	would like to thank Elisabeth Newton and Aaron Dotter for their support and
	for useful disscusion related to the topic of this paper. Additionally, we
	would like to thank Kara Fagerstrom, Aylin Garcia Soto, and Keighley
	Rockcliffe for their useful disscusion related to in this work. We
	acknowledge the support of a NASA grant (No. 80NSSC18K0634). 
\end{acknowledgments}

### 2023-06-30 src/sections/fidanka.tex
Paper updates

\section{fidanka}\label{sec:fidanka}
When fitting isochrones to the data we have four main critera for any method

\begin{itemize}
	\item The method must be robust enough to work along the entire main sequence, turn off, and much of the subgian and red giant branchs.
	\item Any method should consider photometric uncertanity in the fitting process.
	\item The method should be model independent, weighting any n number of populations equally.
	\item The method should be automated and require minimal intervention from the user.
\end{itemize}

While there are many packages which can measure fiducial lines very well
{\color{red}[CITATIONS]}, we do not belive that any of these perfectly match
our use case. Therefore, we elect to develope our own software suite,
\fidanka. \fidanka is a python package designed to automate
much of the process of measuring fiducual lines in CMDs, adhearing to the four
criteria we lay out above. Primary features of \fidanka may be seperated into
two categories: stellar population synthethis and isochrone
optimization/fitting (distance modulus, B-V color excess, and binpary mass
fraction)

\subsection{Fiducual Line Measurment}
\fidanka takes a interative approach to measuring fiducial lines, the first
step of which is to make a ``guess'' as to the fiducial line. This guess will
be used to verticalize the CMD so that further alorithms can work in 1-D
magnitude bins without worrying about weighting issues caused by varying
projections of the evolutionary sequence onto the magnitude axis. This initial
guess is calculated by splitting the CMD into magnitude bins, with uniform
numbers of stars per bin (so that bins are cover a small magnitude range over
densly populated regions of the CMD while covering a much larger mangnitude
range in sparsly populated regions of the CMD, such as the RGB). A unimodal
gaussian distribution is then fit to the color distribution of each bin, and
the resulting mean color is used as thec initial fiducial line guess. This
rough fiducual line will approximatly trace the area of highest density. We
subtract the color of this fiducual line from that of each star to verticalize
the CMD.

If \fidanka were to simply apply the same algorithm to the verticalized CMD
then the resulting fiducial line would simply be a re-extraction of the initial
fiducial line. To avoid this, we take a more robust, number density based
approach, which considers the distribution of stars in both color and magnitude
space simultaniously. For each star in the CMD we first using a
\texttt{introselect} partitioning algorithm to select the 50th nearest stars.
To account for the case where the star is at an extreme edge of the CMD, those
50 stars include the star itself (such that we really select 49 stars + 1). We
use \texttt{qhull}\footnote{https://www.qhull.com}\citep{Barber1996, } to
calculate the convex hull of those 50 points. The number density at each star
then is defined as 50/(area of convex hull). Because we use a fixed number of
points per star, and a partitioning algorithm as opposed to a sorting
algorithm, this method scales like $\mathcal{O}(n)$, where n is the number of
stars in the CMD. It also intrinsically weights the density of of each star
equally as the counting statistics of per convex are uniform. We are left with a
CMD where each star has a defined number density {\color{red}[FIGURE]}.

We can adapt this density map method to consider photometric uncertanties by
adopting a simple monte carlo approach. For each star in the CMD we draw a
random magnitude shift from a gaussian distribution with a standard deviation
equal to the photometric uncertanty of the star in each filter. We then shift
each star by this random amount and calculate the density map as before. We
repeat this process $m$ times, and then take the median density at each point
in the CMD {\color{red}[FIGURE]}. This method will result in noisy density
peaks with charectaristic width less than that of the photometric noise
smoothing out; while wider peaks will remain.

\fidanka can now exploit this density map to fit a better fiducual line to the
data, as the densiy map is far more robust to outlies. There are multiple
algorithms we impliment to fit the fiducial line to the color density profile
in each magnitude bin; they are explained in more detail in the \fidanka
documentation. However, of most relevance here is the A* path finding heuristic
over peaks. Peaks are extracted from the The color-density profile in at
density bin {\color{red}[FIGURE]}. The A* heuristic is then used to find an
optimal path (though not neccesarily {\em the} optimal path) through the peaks
{\color{red}[FIGURE]}. The resulting path is then used as the fiducial line
{\color{red}[FIGURE]}. The heursitc considers the slope of the path from one
bin to the next and simultaniously the change in density from one bin to the
next. This allows the path to be robust to both steep slopes and low density
regions. Due to the heuristic the fiducial line path  is not neccesarily the
vertical line defined by the initial fiducial line guess and will instead be a
more optimized/better guess.

This method of fiducial line extraction is very effective at tracing the
meandian ridge line of the overall CMD; however, it struggles to discriminate
between multiple populations. The density variations from one population to the
next are often too subtle. Moreover, when sampling the density. Moreover, the
spacing between main sequence populations may be of a similar order to the
photmetric uncertantities, and therefor the indivudual populations may smear
into one and other. For these reasons, \fidanka does not attempt to extract
multiple unique fiducuial lines for each population; instead, it measures the
width of the overall sequence. Width measurment again makes use of the
color-density pofile, selecting defining the width as the difference in color
between the 5th and 95th percentile of the density {\color{red}[FIGURE]}.

\subsection{Stellar Population Synthethis}
In addition to measuring fiducial lines, \fidanka also includes a stellar
population synthethis module. This module is used to generate synthetic CMDs
from a given set of isochrones. This is of primary importance for binary
population modelling. The module is also used to generate synthetic CMDs for
the purpose of testing the fiducial line extraction algorithms against priors.

\fidanka uses MIST formated isochrones {\color{red}[CITATION]} as input along
with distance modulus, B-V color excess, binary mass fraction, and bolometric
corrections. An arbitrarily large number of isochrones may be used to define an
arbitrary number of populations. Synthetic stars are samples from each
isochroner based on a definiable probability (for example it is belived that
$\sim90\%$ of stars in globular clusters are younger population
{\color{red}[CITATION]). Based on the metallicity, $\mu$, and E(B-V) of each
isochrone, bolometric corrections are taken from bolometric correction
tables. Where bolometric correction tables do not include exact metallicities
or extinctions a linear interpolation is preformed between the two bounding
values. {\color{red}[FIGURE]} shows an example of a synthetic CMD generated
from a set of 2 NGC 2808 isochrones as well as a comparison between those isochrones and
the measured fiducual line of the synthetic population.

>>>>

\section{fidanka}\label{sec:fidanka}
When fitting isochrones to the data we have four main criteria for any method

\begin{itemize}
	\item The method must be robust enough to work along the entire main sequence, turn off, and much of the subgiant and red giant branchs.
	\item Any method should consider photometric uncertainty in the fitting process.
	\item The method should be model independent, weighting any n number of populations equally.
	\item The method should be automated and require minimal intervention from the user.
\end{itemize}

We do not believe that any currently available software is a match for
our use case. Therefore, we elect to develop our own software suite, \fidanka.
\fidanka is a python package designed to automate much of the process of
measuring fiducial lines in CMDs, adhering to the four criteria we lay out
above. Primary features of \fidanka may be separated into three primary
categories: fiducial line measurement, stellar population synthesise, and
isochrone optimization/fitting. Additionally, there are utility functions which
are detailed in the \fidanka documentation.

\subsection{Fiducial Line Measurement}
\fidanka takes a iterative approach to measuring fiducial lines, the first step
of which is to make a ``guess'' as to the fiducial line. . This initial guess
is calculated by splitting the CMD into magnitude bins, with uniform numbers of
stars per bin (so that bins are cover a small magnitude range over densely
populated regions of the CMD while covering a much larger magnitude range in
sparsely populated regions of the CMD, such as the RGB). A unimodal Gaussian
distribution is then fit to the color distribution of each bin, and the
resulting mean color is used as the initial fiducial line guess. This rough
fiducial line will approximately trace the area of highest density. The initial
guess will be used to verticalze the CMD so that further algorithms can work in
1-D magnitude bins without worrying about weighting issues caused by varying
projections of the evolutionary sequence onto the magnitude axis.
Verticalization is preformed taking the difference between the guess fiducial
line and the color of each star in the CMD.

If \fidanka were to simply apply the same algorithm to the verticalized CMD
then the resulting fiducial line would likely be a re-extraction of the initial
fiducial line guess. To avoid this, we take a more robust, number density based
approach, which considers the distribution of stars in both color and magnitude
space simultaneously. For each star in the CMD we first using a
\texttt{introselect} partitioning algorithm to select the 50th nearest stars.
To account for the case where the star is at an extreme edge of the CMD, those
50 stars include the star itself (such that we really select 49 stars + 1). We
use \texttt{qhull}\footnote{https://www.qhull.com}\citep{Barber1996, } to
calculate the convex hull of those 50 points. The number density at each star
then is defined as $50/A_{hull}$, where $A_{hull}$ is the area of the convex
hull. Because we use a fixed number of points per star, and a partitioning
algorithm as opposed to a sorting algorithm, this method scales like
$\mathcal{O}(n)$, where n is the number of stars in the CMD. This method also
intrinsically weights the density of of each star equally as the counting
statistics per bin are uniform. We are left with a CMD where each star
has a defined number density (Figure \ref{fig:densityMapDemo}).

\begin{figure*}
	\centering
	\includegraphics[width=0.9\textwidth]{src/figures/notebookFigures/DensityMapDemo.png}
	\label{fig:densityMapDemo}
	\caption{Density map demo showing density estimate over different parts of
	the evolutionary sequence. The left panel shows the density map over the
	entire evolutionary sequence, while the middle panel shows the density map
	over the main sequence and the right most panel shows the density map over
	the RGB. Figures in the top row are the raw CMD, while figures in the
	bottom row are colored by the density map.}
\end{figure*}

\fidanka can now exploit this density map to fit a better fiducial line to the
data, as the density map is far more robust to outliers. There are multiple
algorithms we implement to fit the fiducial line to the color-density profile
in each magnitude bin (Figure \ref{fig:densityBinsDemo}); they are explained in more detail in the \fidanka
documentation. However, of most relevance here is the Bayesian Gaussian Mixture
Modeling (BGMM) method. BGMM is a clustering algorithm which, for some fixed
number of n-dimensional Gaussian distributions, $K$, determines the mean, covariance, and
mixing probability (somewhat analogous to amplitude) of each $k^{th}$
distribution, such that the local lower bound of the evidence of each star
belonging strongly to a single distribution is maximized.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{src/figures/notebookFigures/DensityBinsDemo.png}
	\label{fig:densityBinsDemo}
	\caption{CMD where points are colored by density. Lines show the
	density-color profile in each magnitude bin. In this figure adaptive
	binning targeted 1000 stars per bin}
\end{figure}

Maximization is preformed using the Dirichlet process, which is a
non-parametric Bayesian method of determining the number of Gaussian distributions, $K$,
which best fit the data {\color{red}[CITATION]}. Use of the Dirichlet process
allows for dynamic variation in the number of inferred populations from
magnitude bin to magnitude bin. Specifically, populations are clearly visually
separated from the lower main sequence through the turn off; however, at the
turn off and throughout much of the subgiant branch, the two visible
populations overlap due to their extremely similar ages ({\color{red}
Citation}). The Dirichlet process allows for the BGMM method to infer a single
population in these regions, while inferring two populations in regions where
they are clearly separated. More generally, the use of the Dirichlet process
removes the need for a prior on the exact number of populations to fit. Rather,
the user specifies a upper bound on the number of populations within the
cluster. An example bin (F814W = 20.6) is shown in Figure \ref{fig:BGMMDist}.

\begin{figure*}
	\centering
	\includegraphics[width=0.9\textwidth]{Notebooks/Figures/BGMMMixingBin.pdf}
	\caption{Example of BGMM fit to a magnitude bin. The grey line shows the
	underlying color-density profile, while the black dashed-line shows the
	joint distribution of each BGMM component. The solid black lines show the
	two selected components.}
	\label{fig:BGMMDist}
\end{figure*}

\fidanka's BGMM method first breaks down the verticalized CMD into magnitude
bins with uniform numbers of stars per bin (here we adopt 250). Any stars left
over are placed into the final bin. For each bin a BGMM model with a maximum of
5 populations is fit to the color density profile. The number of populations is
then inferred from the weighting parameter (the mixing probability) of each
population. If the weighting parameter of any $k^{th}$ components less than
{\color{blue}0.05}, then that component is considered to be spurious and
removed. Additionally, if the number of populations in the bin above and the
bin below are the same, then the number of populations in the current bin is
forced to be the same as the number of populations in the bin above. Finally,
the initial guess fiducial line is added back to the BGMM inferred line. Figure
\ref{fig:vertFit} shows the resulting fiducial line(s) in each magnitude bin
for both a verticalized CMD and a non verticalized CMD.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{Notebooks/Figures/vertFit.png}
	\caption{CMD where points are colored by density. Line trace the infered 
	fiducial line(s) in each magnitude bin.}
	\label{fig:vertFit}
\end{figure}

This method of fiducial line extraction effectively discriminated between
multiple populations long the main sequence and RGB of a cluster, while
simultaneously allowing for the presence of a single population along the MSTO
and subgiant branch.

We can adapt this density map based BGMM method to consider photometric
uncertainties by adopting a simple Monte Carlo approach. Instead of measuring
the fiducial line(s) a single time, \fidanka can measure the fiducial line(s)
many times, resampling the data with replacement each time. For each resampling
\fidanka adds a random offset to each filter based on the photometric
uncertainties of each star. From these $n$ measurments the mean fiducial line
for each sequence can be identified along with upper and lower bound confidence
intervals in each magnitude bin.

\subsection{Stellar Population Synthesis}
In addition to measuring fiducial lines, \fidanka also includes a stellar
population synthesise module. This module is used to generate synthetic CMDs
from a given set of isochrones. This is of primary importance for binary
population modelling. The module is also used to generate synthetic CMDs for
the purpose of testing the fiducial line extraction algorithms against priors.

\fidanka uses MIST formatted isochrones {\color{red}[CITATION]} as input along
with distance modulus, B-V color excess, binary mass fraction, and bolometric
corrections. An arbitrarily large number of isochrones may be used to define an
arbitrary number of populations. Synthetic stars are samples from each
isochrone based on a definable probability (for example it is believed that
$\sim90\%$ of stars in globular clusters are younger population
{\color{red}[CITATION]}). Based on the metallicity, $\mu$, and E(B-V) of each
isochrone, bolometric corrections are taken from bolometric correction tables.
Where bolometric correction tables do not include exact metallicities or
extinctions a linear interpolation is preformed between the two bounding
values. {\color{red}[FIGURE]} shows an example of a synthetic CMD generated
from a set of 2 NGC 2808 isochrones as well as a comparison between those
isochrones and the measured fiducial line of the synthetic population.

\subsection{Isochrone Optimization}
The optimization routines in \fidanka will find the best fit distance modulus,
B-V color excess, and binary number fraction for a given set of isochrones. If
a single isochrone is provided then the optimization is done by minimizing the
$\chi^2$ of the perpendicular distances between an isochrone and a fiducial
line. If multiple isochrones are provided then those isochrones are first used
to run stellar population synthesis and generate a synthetic CMD. The
optimization is then done by minimizing the $\chi^2$ of both the perpendicular
distances between and widths of the observed fiducial line and the fiducial
line of the synthetic CMD.

\subsection{Fidanka Testing}
In order to validate fidanka we have run a series of tests on synthetic CMDs
generated from NGC 2808 isochrones. For each test we use fidanka to both build
a synthetic population and then measure the fiducial line of that population.
The measured fiducial lines(s) are then compared to the underlying isochrones
used.

### 2023-06-30 src/sections/intro.tex
Paper updates

\section{Introduction}\label{sec:Intro}
Globular clusters (GCs) are among the oldest observ- able objects in the
universe \citep{Pen11}. They are characterized by high densities with typical
half-light radii of $\le$10 pc \citep{Vanderburg2010}, and typical masses
ranging from $10^{4}$--$10^{5}$ M$_{\odot}$ \citep{Bro06} --- though some GCs
are signifigantly larger than these typical values {\color{red}EXAMPLE}. GCs
provide a unique way to probe stellar evolution \citep{Bau03}, galaxy formation
models \citep{Boy18,Kra05}, and dark matter halo struc- ture \citep{Hud18}.
{\color{red} BRING IN SOME MORE RECENT CITATIONS.}

Whereas, people have have often tried to categorized objects as GCs through
relations between half-light radius, density, and surface brightness profile,
in fact many objects which are generally thought of as GCs don't cleanly fit
into these cuts {\color{red} EXAMPLE + CITATION}. Consequently,
\citet{Carretta2010} proposed a definition of GC based on observed chemical
inhomogeneities in their stellar populations. The modern understanding of GCs
then is not simply one of a dense cluster of stars which may have chemical
inhomogeneities and multiple populations; rather, it is one where those
chemical inhomogeneities and multiple populations themselves are the defining
element of a GC.

@@

Formation channels for these multiple populations remain a point of debate
among astronomers. Most proposed formation channels consist of some older,
more massive, population of stars polluting the pristine cluster media before a
second population forms, now enriched in heavier elements which they themselves could
not have generated \citep[for a detailed review see ][]{Gratton2012}. The four
primary candidates for these polluters are asymptotic giant branch stars
\citep[AGBs,][]{Ventura2001,DErcole2010}, fast rotating massive stars
\citep[FRMSs,][]{Decressin2007}, super massive stars
\citep[SMSs,][]{Denissenkov2014}, and massive interacing binaries
\citep[MIBs,][]{deMink2009, Bastian2018}.

Hot hydrogen burning (proton capture), material transport to the surface, and
material ejection into the intra-cluster media are features of each of these
models and consequently they can all be made to {\it qualitatively} agree with
the observed elemental abundances. However, none of the standard models can
currently account for all specific abundances \citep{Gratton2012}. AGB and FRMS
models are the most promising; however, both models have difficulty reproducing
severe O depletion \citep{Ventura2009,Decressin2007}. Moreover, AGB and FRMS
models require signifigant mass loss ($\sim 90\%$) between cluster formation
and the current epoch --- implying that a signifigant fraction of halo stars
formed in GCs \citep{Renzini2008,DErcole2008,Bastian2015}.

In addition to the light-element anti-correlations observed it is also known
that younger populations are signifigantly enhanced in Helium
\citep{Piotto2007, Piotto2015, Latour2019}. Depending on the cluster, Helium
mass fractions as high as $Y=0.4$ have been inferred \citep[e.g][]{Milone2015}.
However, due to the relatively high and tight temperature range of partial
ionization for He it cannot be observed in globular clusters; consequently, the
evidence for enhanced He in GCs originates from comparison of theoretical
stellar isochrones to the observed color-magnitude-diagrams of globular
clusters. Therefore, a careful handling of chemistry is essential when modeling
with the aim of discriminating between MPs; yet, only a very limited number of
GCs have yet been studied with chemically self-consistent (structure and
atmosphere) isochrones \citep[e.g.][NGC 6752]{Dotter2015}. Here we present new,
chemically self-consistent modeling og the two extreme populations of NGC 2808
identified by \citet{Milone2015}, A and E.

>>>>

\section{Introduction}\label{sec:Intro}
Globular clusters (GCs) are among the oldest observable objects in the
universe \citep{Pen11}. They are characterized by high densities with typical
half-light radii of $\le$10 pc \citep{Vanderburg2010}, and typical masses
ranging from $10^{4}$--$10^{5}$ M$_{\odot}$ \citep{Bro06} --- though some GCs
are significantly larger than these typical values {\color{red}EXAMPLE}. GCs
provide a unique way to probe stellar evolution \citep{Bau03}, galaxy formation
models \citep{Boy18,Kra05}, and dark matter halo structure \citep{Hud18}.
{\color{red} BRING IN SOME MORE RECENT CITATIONS.}

The traditional view of Globular Clusters was, for a long time, that they
consisted of a single stellar population (SSP, in some publications this is
referred to as a Simple Stellar Population). This view was supported by
spectroscopically uniform heavy element abundances \citep{Carretta2009}
{\color{red}[CHECK]}, and the lack of evidence for multiple stellar populations
(MPs) in the color-magnitude diagrams of GCs \citep{Milone2017}. However, in
the last two decades, it has been definitively shown that most if not all Milky
Way GCs have MPs \citep{Piotto2015}. The lack of photometric evidence for MPs
can be attributed to the short color throw available to ground based
photometric surveys \citep{Milone20Here17}. While MPs are chemically distinct
from one another, that distinction is most prominent when observing with $U$
and $B$ filters \citep{Sbordone2011}.

The prevalence of multiple populations in GCs is so distinct that the proposed
definitions for what constitutes a globular cluster now often center the
existence of MPs. Whereas, people have have often tried to categorized objects
as GCs through relations between half-light radius, density, and surface
brightness profile, in fact many objects which are generally thought of as GCs
don't cleanly fit into these cuts {\color{red} EXAMPLE + CITATION}.
Consequently, \citet{Carretta2010} proposed a definition of GC based on
observed chemical inhomogeneities in their stellar populations. The modern
understanding of GCs then is not simply one of a dense cluster of stars which
may have chemical inhomogeneities and multiple populations; rather, it is one
where those chemical inhomogeneities and multiple populations themselves are
the defining element of a GC.

@@

Formation channels for these multiple populations remain a point of debate
among astronomers. Most proposed formation channels consist of some older,
more massive, population of stars polluting the pristine cluster media before a
second population forms, now enriched in heavier elements which they themselves could
not have generated \citep[for a detailed review see ][]{Gratton2012}. The four
primary candidates for these polluters are asymptotic giant branch stars
\citep[AGBs,][]{Ventura2001,DErcole2010}, fast rotating massive stars
\citep[FRMSs,][]{Decressin2007}, super massive stars
\citep[SMSs,][]{Denissenkov2014}, and massive interacting binaries
\citep[MIBs,][]{deMink2009, Bastian2018}.

Hot hydrogen burning (proton capture), material transport to the surface, and
material ejection into the intra-cluster media are features of each of these
models and consequently they can all be made to {\it qualitatively} agree with
the observed elemental abundances. However, none of the standard models can
currently account for all specific abundances \citep{Gratton2012}. AGB and FRMS
models are the most promising; however, both models have difficulty reproducing
severe O depletion \citep{Ventura2009,Decressin2007}. Moreover, AGB and FRMS
models require significant mass loss ($\sim 90\%$) between cluster formation
and the current epoch --- implying that a significant fraction of halo stars
formed in GCs \citep{Renzini2008,DErcole2008,Bastian2015}.

In addition to the light-element anti-correlations observed it is also known
that younger populations are significantly enhanced in Helium
\citep{Piotto2007, Piotto2015, Latour2019}. Depending on the cluster, Helium
mass fractions as high as $Y=0.4$ have been inferred \citep[e.g][]{Milone2015}.
However, due to the relatively high and tight temperature range of partial
ionization for He it cannot be observed in globular clusters; consequently, the
evidence for enhanced He in GCs originates from comparison of theoretical
stellar isochrones to the observed color-magnitude-diagrams of globular
clusters. Therefore, a careful handling of chemistry is essential when modeling
with the aim of discriminating between MPs; yet, only a very limited number of
GCs have yet been studied with chemically self-consistent (structure and
atmosphere) isochrones \citep[e.g.][NGC 6752]{Dotter2015}.

{\color{blue} NGC 2808 is the prototype globular cluster to host Multiple
Populations.} Various studies since 2007 have identified that it may host
anywhere from 2-5 stellar populations. These populations have been identified
both spectroscopically \citep[i.e.][]{} and photometrically \citep[i.e.][]{}.
Note that recent work \citep{Valle2022} calls into question the statistical significance
of the detections of more than 2 populations in the spectroscopic data. Here we
present new, chemically self-consistent modeling of the two extreme populations
of NGC 2808 identified by \citet{Milone2015}, populations A and E.
{\color{blue} Additionally, we present a likelihood analysis of the photometric
data of NGC 2808 to determine the number of populations present in the cluster.}

### 2023-06-30 src/sections/modeling.tex
Paper updates

\section{Modeling}\label{sec:modeling}
One key element of NGC 2808 modeling is the incorporation of new atmospheric
models, generated from the \texttt{MARCS} grid of model atmospheres \citep{Plez2008},
which match interior elemental abundances. \texttt{MARCS} provides one-dimensional,
hydrostatic, plane-parallel and spherical LTE atmopsheric models
\citep{Gustafsson2008}. Members of our collaboration have generated atmospheric 
models for populations A and E.  Integration of these new model atmospheres
into DSEP is ongoing.

For similar reasons as discussed in \S\ref{sec:p1} we conduct this research
with OPLIB high-temperature opacity tables as opposed to OPAL tables. We will
also generate low temperature opacity tables using the \texttt{MARCS}.
Moreover, we confirm that the atmosphere and structure meet in an optically
thick region of the star by shifting the atmospheric fitting point from an optical
depth of $\tau = 2/3$ (used by DSEP currently for \texttt{PHOENIX} model atmospheres) to
some higher $\tau$. We will experiment to identify the best optical depth to
fit at..

These population have been studied in depth by Feiden and their chemical
compositions were determined in \citet{Milone2015} (see Table 2 in that paper).
While we cannot yet evolve DSEP models with these new boundary conditions, we
can make a first pass investigation of the affect of OPLIB opacities (Figure
\ref{fig:NGC2808ISO}). Note how the models generated using OPLIB opacity tables
have a systematically lower luminosity. This discrepancy is consistent with
the overall lower opacities of the OPLIB tables.

The isochrones generally used to infer the degree of helium enhancements assume that
convection operates in the same manner in metal-poor stars as it does in the
Sun. However, observations from \textit{Kepler} of metal-poor red giants
\citep{Bonaca2012, tayar2017correlation}, in concert with interferometric
radius determination of the metal-poor sub-giant HD 140283
\citep{creevey2015benchmark}, have shown that the efficiency of convection
changes with iron content. As the final portion of our work to more carefully
handle a star's chemistry, we will modify DSEP to capture this variation in
convective efficiency.

>>>>

\section{Stellar Models}\label{sec:modeling}
We use the Dartmouth Stellar Evolution Program \citep[DSEP, ][]{Dotter2008} to
generate stellar models. DSEP is a well-tested, one-dimensional stellar
evolution code which includes a mixing length model of convection,
gravitational settling, and diffusion.

We use DSEP to evolve stellar models ranging in mass from 0.3 to 2.0 solar
masses from the zero-age main sequence (ZAMS) to the tip of the red giant
branch. Below 0.7 $M_{\odot}$ we evolve a model every 0.03 $M_{\odot}$ and
above 0.7 $M_{\odot}$ we evolve a model every 0.5 $M_{\odot}$. Additionally, we
evolve models over a grid of mixing length parameters, $\alpha_{MLT}$, from
$\alpha_{MLT} = 1.0$ to $\alpha_{MLT} = 2.0$ in steps of 0.1. In addition to the
mixing length grid the evolved grid of models also has dimensions population (A
or E) and helium abundance ($Y$). Each model is evolved in DSEP's ``high
resolution'' mode and had a maximum allowed time step of 50 Myr.

For each combination of population, $Y$, and $\alpha_{MLT}$ we use the
isochrone generation code first presented in \citet{Dotter2016} to generate a
grid of isochrones. The isochrone generation code identified equivalent
evolutionary points (EEPs) over a series of masses and interpolates between
them. {\color{blue} The grid of isochrones generated for this work is avalible
as a digital supplement to this paper.}
